Strong coupling between scales in a multi-scalar model of urban dynamics

### 2024-10-01 Docs/Papers/MesoBenchmark/paper/paper.tex
revision_multiscale
revisions
.
bib
papers

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
%\documentclass[10pt]{article}
%\usepackage[margin=2cm]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
\usepackage{setspace} 
\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{A comparison of simple models for urban morphogenesis} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juste Raimbault\textsuperscript{1,2,3,*}
\bigskip\\
\textsuperscript{1} Center for Advanced Spatial Analysis, University College London
\\
\textsuperscript{2} UPS CNRS 3611 ISC-PIF
\\
\textsuperscript{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
\bigskip\\
* juste.raimbault@polytechnique.edu

\end{flushleft}

% Please keep the abstract below 300 words
\section*{Abstract}
\vspace{1cm}
The spatial distribution of population and activities within urban areas, or urban form at the mesoscopic scale, is the outcome of multiple antagonist processes. We propose in this paper to benchmark different models of urban morphogenesis, to systematically compare the urban forms they can produce. Different types of approaches are included, such as a reaction-diffusion model, a gravity-based model, and correlated percolation. Applying a diversity search algorithm, we estimate the feasible space of each model within a space of urban form indicators, in comparison of empirical values for worldwide urban areas. We find a complementarity of the different types of processes, advocating for a plurality of urban models.

% urban morphogenesis; reaction-diffusion; correlated percolation; gravity model; novelty search

\linenumbers

@@

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	%\includegraphics[width=\linewidth]{figures/Fig1.png}
	\vspace{2cm}
	\caption{{\bf Maps of urban morphology indicators for worldwide urban areas.} (Top) Rank-size hierarchy; (Bottom) Moran spatial autocorrelation index. Area of circles gives population.\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%

@@

%\section*{Supporting information}

%\section*{Acknowledgments}

\nolinenumbers

>>>>

\documentclass[referee,lineno,pdflatex,sn-apa]{sn-jnl}

\jyear{2024}%

\raggedbottom

\begin{document}

\title[A comparison of simple models for urban morphogenesis]{A comparison of simple models for urban morphogenesis}

\author{}
%\author*[1,2,3,4]{\pfx{Dr} \fnm{Juste} \sur{Raimbault} }\email{juste.raimbault@ign.fr}% \dgr{PhD} % orcid 0000-0003-0768-9480

%\affil*[1]{\orgdiv{LASTIG}, \orgname{Univ. Gustave Eiffel, IGN-ENSG}} %, \orgaddress{\street{73 avenue de Paris}, \city{Saint-Mand{\'e}}, \postcode{93160}, \country{France}}
%\affil[2]{\orgdiv{Center for Advanced Spatial Analysis}, \orgname{UCL}} % , \orgaddress{\street{90 Tottenham Court Road}, \city{London}, \postcode{}, \country{UK}}
%\affil[3]{\orgdiv{UPS CNRS 3611 ISC-PIF}, \orgname{CNRS}} % , \orgaddress{\street{Place Nationale}, \city{Paris}, \postcode{75013}, \country{France}}
%\affil[4]{\orgdiv{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s}, \orgname{Universit{\'e} Paris 1}} % , \orgaddress{\street{Place Nationale}, \city{Paris}, \postcode{75013}, \country{France}}

% Target journal : Computational Urban Science https://link.springer.com/journal/43762/submission-guidelines

% TODO Reviews Plos One

\abstract{The spatial distribution of population and activities within urban areas, or urban form at the mesoscopic scale, is the outcome of multiple antagonist processes. We propose in this paper to benchmark different models of urban morphogenesis, to systematically compare the urban forms they can produce. Different types of approaches are included, such as a reaction-diffusion model, a gravity-based model, and correlated percolation. Applying a diversity search algorithm, we estimate the feasible space of each model within a space of urban form indicators, in comparison of empirical values for worldwide urban areas. We find a complementarity of the different types of processes, advocating for a plurality of urban models.}

\keywords{Urban morphogenesis; Reaction-diffusion; Correlated percolation; Gravity model; Novelty search}

\maketitle

Batty, M. (2001). Polynucleated urban landscapes. Urban studies, 38(4), 635-655. model to include?

\cite{xu2021emergence}

\cite{strano2021agglomeration}

\cite{rybski2021comparing}

\cite{Rybski2021Modeling} coupling inside model - more to benchmark - multimodeling

\cite{bithellsystematic} benchmark abm covid

@@

\cite{Nicholse2100769118} reproducibility - benchmarks - sequence of studies

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	%\includegraphics[width=\linewidth]{figures/Fig1.png}
	\vspace{2cm}
	\caption{{\bf Maps of urban morphology indicators for worldwide urban areas.} (Top) Rank-size hierarchy; (Bottom) Moran spatial autocorrelation index. Area of circles gives population.\label{fig:fig1}}
% TODO redo maps with open resources https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
\end{figure}
%%%%%%%%%%%%%%

@@

### 2024-10-01 Docs/Papers/Multiscale/paper/paper.tex
.
papers

\jyear{2022}%

@@

@@

\abstract{Urban evolution processes occur at different scales, with intricate interactions between levels and relatively distinct type of processes. To what extent actual urban dynamics include an actual strong coupling between scales, in the sense of both top-down and bottom-up feedbacks, remains an open issue with important practical implications for the sustainable management of territories. We introduce in this paper a multi-scalar simulation model of urban growth, coupling a system of cities interaction model at the macroscopic scale with morphogenesis models for the evolution of urban form at the scale of metropolitan areas. Strong coupling between scales is achieved through an update of model parameters at each scale depending on trajectories at the other scale. The model is applied and explored on synthetic systems of cities. Simulation results show a non-trivial effect of the strong coupling. As a consequence, an optimal action on policy parameters such as containing urban sprawl is shifted. We also run a multi-objective optimization algorithm on the model, showing showing that compromise between scales are captured. Our approach opens new research directions towards more operational urban dynamics models including a strong feedback between scales.}

@@

Modelling urban growth and more generally the dynamics of urban systems is central to the design of sustainable territorial policies, enabling the understanding of past urbanisation processes and the anticipation of future urban trajectories. The design of sustainable future cities requires an historical knowledge of how past cities came to be and evolved \citep{batty2018inventing}. Several models have been developed at different scales and integrating multiple urban systems dimensions. Some typical examples are models of land-use change at a mesoscopic scale \citep{iacono2008models}, or systems of cities models at a macroscopic scale \citep{pumain2017urban}.

@@

In disciplines close to urban modelling, methods have been developed for multi-scale models. For example in spatial epidemiology, \cite{banos2015importance} combines agent-based modelling for local diffusion dynamics with differential equations at the population scale. The NetLogo software for agent-based modelling includes a specific extension for multi-scale modelling \cite{hjorth2020levelspace}. Multi-scale models have also been used for the simulation of crowd dynamics \cite{crociani2016multi}. The study of traffic is also made more accurate by coupling macroscopic and microscopic models \cite{boulet2020coupling}. The management of ecosystems  requires integrating across actors and scales \cite{belem2013organizational}. These non-exhaustive illustrations highlight how the integration of scales is a crucial feature and issue in the understanding of complex systems \cite{chavalarias2009french}.

@@

Model behaviour is characterised using aggregated indicators at each scale considered. At the  macroscopic scale, we follow \cite{raimbault2020unveiling} and consider distributions of population, accessibilities, and centralities, all three summarised by their respective average, hierarchy, and entropy. At the mesoscopic scale, urban form is captured following the indicators used by \cite{raimbault2018calibration}, i.e. Moran index, average distance between individuals, hierarchy of population distribution, and entropy of population distribution.

@@

Model implementation has to meet some performance constraints: $N$ mesoscopic morphogenesis models must be simulated in parallel at each macroscopic time step. On the contrary, macroscopic interactions are efficient to implement as they are based on matrices computations. The model is implemented in \texttt{scala} and integrated into a broader library for spatial sensitivity analysis and dynamical models for settlement systems, namely the \texttt{spatialdata} library \cite{raimbault2020scala}. This library in particular already includes implementations of the models used in \cite{raimbault2020indirect} and \cite{raimbault2018calibration}, but also of other urban interaction models such as \cite{favaro2011gibrat}. The relative large number of parameters and output indicators imposes the use of appropriate model exploration methods to get a grasp of model behavior. Therefore, the parameter space is explored with the OpenMOLE model exploration software \citep{reuillon2013openmole}. Standalone source code for the model is available on an open git repository as \cite{urbangrowthmodel}. An other open repository for the analysis of results is also available as \cite{urbangrowth}. Simulation data used in this paper is available on the dataverse repository as \cite{dataverse}.

>>>>

\jyear{2024}%

@@

% Target: Urban Informatics
% https://www.springer.com/journal/44212/submission-guidelines
% check Response_to_Reviewers submission IJGI

@@

\abstract{Urban dynamics occur at different scales, with intricate interactions between levels and relatively distinct type of processes. To what extent observed urban trajectories include an actual strong coupling between scales, in the sense of both top-down and bottom-up feedbacks, remains an open issue with important practical implications for the sustainable management of territories. We introduce in this paper a multi-scalar simulation model of urban growth, coupling a system of cities interaction model at the macroscopic scale with morphogenesis models for the evolution of urban form at the scale of metropolitan areas. Strong coupling between scales is achieved through an update of model parameters at each scale depending on trajectories at the other scale. The model is applied and explored on synthetic systems of cities. Simulation results show a non-trivial effect of the strong coupling. As a consequence, an optimal action on policy parameters such as containing urban sprawl is shifted. We also run a multi-objective optimization algorithm on the model, showing showing that compromise between scales are captured. Our approach opens new research directions towards more operational urban dynamics models including a strong feedback between scales.}

@@

Modelling urban growth and more generally the dynamics of urban systems is central to the design of sustainable territorial policies, enabling the understanding of past urbanisation processes and the anticipation of future urban trajectories \citep{pumain2017urban}. The design of sustainable future cities requires an historical knowledge of how past cities came to be and evolved \citep{batty2018inventing}. Several models have been developed at different scales and integrating multiple urban systems dimensions. Some typical examples are models of land-use change at a mesoscopic scale \citep{iacono2008models}, or systems of cities models at a macroscopic scale \citep{raimbault2020empowering}.

@@

In disciplines that are neighbour to urban modelling, methods have been developed for multi-scale models. For example in spatial epidemiology, \cite{banos2015importance} combines agent-based modelling for local diffusion dynamics with differential equations at the population scale. The NetLogo software for agent-based modelling includes a specific extension for multi-scale modelling \cite{hjorth2020levelspace}. Multi-scale models have also been used for the simulation of crowd dynamics \cite{crociani2016multi}. The study of traffic is also made more accurate by coupling macroscopic and microscopic models \cite{boulet2020coupling}. The management of ecosystems  requires integrating across actors and scales \citep{belem2013organizational}. These non-exhaustive illustrations highlight how the integration of scales is a crucial feature and issue in the understanding of complex systems \citep{chavalarias2009french}.

In other disciplines such as physics or engineering, the methodology and application of multi-scale modelling is much more mature.
% TODO examples physics

@@

Model behaviour is characterised using aggregated indicators at each scale considered. At the  macroscopic scale, we follow \cite{raimbault2020unveiling} and consider distributions of population, accessibilities, and centralities, all three summarised by their respective average, hierarchy, and entropy. At the mesoscopic scale, urban form is captured using the indicators proposed by \cite{raimbault2018calibration}, i.e. Moran index, average distance between individuals, hierarchy of population distribution, and entropy of population distribution.

@@

Model implementation has to meet some performance constraints: $N$ mesoscopic morphogenesis models must be simulated in parallel at each macroscopic time step. On the contrary, macroscopic interactions are efficient to implement as they are based on matrices computations. The model is implemented in \texttt{scala} and integrated into a broader library for spatial sensitivity analysis and dynamical models for settlement systems, namely the \texttt{spatialdata} library \citep{raimbault2020scala}. This library in particular already includes implementations of the models used in \cite{raimbault2020indirect} and \cite{raimbault2018calibration}, but also of other urban interaction models such as \cite{favaro2011gibrat}. The relative large number of parameters and output indicators imposes the use of appropriate model exploration methods to get a grasp of model behavior. Therefore, the parameter space is explored with the OpenMOLE model exploration software \citep{reuillon2013openmole}. Standalone source code for the model is available on an open git repository as \cite{urbangrowthmodel}. An other open repository for the analysis of results is also available as \cite{urbangrowth}. Simulation data used in this paper is available on the dataverse repository as \cite{dataverse}.
% TODO anon repos
